Global patterns of forest autotrophic carbon fluxes

### 2019-09-17 manuscript/manuscript-draft.tex
finished integrating Nobby's comments
There are a few citation lists not yet copied over, but I've added notes to see comments in his .pdf.

@@

Carbon fixation, allocation, and metabolism by trees sets the basis for
energy and material flows in forest ecosystems and defines their
interactions with Earth's changing climate. Forest autotrophic carbon
flux (FACF) influences all organic matter stocks in ecosystems and is
linked to cycling of energy, water, and nutrients (\textbf{REFS}).
Forest productivity sets the energy available to heterotrophic
communities, in turn influencing their abundance (\textbf{REFS}) and
possibly diversity (\textbf{REFS})--but probably not individual
metabolic rate (\textbf{Anderson \& Jetz 2005, but check more recent
literature}). On the global scale, forests play a critical role in
regulating atmospheric CO\_2 and climate (\textbf{Bonan}). The total
amount of CO\_2 cycling through Earth's forests each year (\emph{i.e.},
total gross primary production, \(GPP\)) is more than five times
anthropogenic fossil fuel emissions, and net sequestration
(\textasciitilde{}2.4 GtCyr\textsuperscript{-1}) offsetting roughly 30\%
of these emissions \citep{pan_large_2011}. As atmospheric carbon dioxide
levels continue to rise, with consequences for global climate, there is
increasing recognition that proper protection and management of forest
resources will have an important role to play in mitigating climate
change (** REFS- e.g., IPCC 1.5C and land reports**). Given the
importance of understanding broad-scale patterns in forest autotrophic
carbon flux (FACF), there have been many meta-analyses on the theme.
However, our ability to draw general macroscopic conclusions regarding
global variation in multiple productivity variables with respect to
climate has been limited in that these analyses often mix forests that
vary in stand age, disturbance history, and/or management status; do not
always sufficiently parse related variables (\emph{e.g.}, combining net
primary productivity records with and without belowground components);
and typically consider only one or a few variables at a time. The recent
development of a global forest carbon database synthesizing multiple
variables and including records of stand history (ForC;
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
opens up the possibility for a standardized analysis of global scale
variation in multiple components of forest productivity and the
principle climatic drivers of these patterns.

On a global scale, the productivity of forests varies with latitude,
showing a general trend of decreasing productivity with latitude
\citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
productivity is lowest in the boreal regions, and increases into the
temperate regions
\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
However, evidence is inconclusive on whether productivity continues to
increase into the tropics, or whether it plateaus in temperate regions.
Evidence for this is further complicated by the fact that different
studies use different measures of productivity to explore these
relationships. For example, modelling of global terrestrial ecosystem
gross primary productivity (GPP) through upscaling and calibration of
eddy flux measurements indicates that GPP peaks in tropical forests
\citep{beer_terrestrial_2010, jung_global_2011};\href{https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729}{Bagdley
et al. 2019}). This is corroborated by analysis of site-level GPP
measurements, which appear to reach their highest levels in tropical
forests \citep{luyssaert_co_2007}. In contrast, there is evidence that
the highest values of net primary productivity (NPP) may be found in
temperate forests \citep{luyssaert_co_2007, huston_global_2009},
although other studies find NPP is highest in the tropics, showing a
decrease with latitude \citep{simova_enigma_2017}. Other studies have
chosen to focus exclusively on above-ground net primary productivity
(ANPP), finding evidence of a weak negative relationship between ANPP
and latitude \citep{huston_global_2009, gillman_latitude_2015}. An
understanding of the global patterns of productivity is important;
however elucidating the drivers of these patterns will be even more
valuable in expanding our understanding of global carbon cycling.
Primary productivity can be influenced by many factors, which often act
across a range of scales, and may show interactive effects with each
other. On a local scale, stand age
\citep{litton_carbon_2007, gillman_latitude_2015}, management
\citep{simova_enigma_2017}; nutrient availability
\citep{aragao_above-_2009}; and altitude
\citep{girardin_net_2010, malhi_variation_2017} all impact forest
productivity. On a global scale, we expect that productivity is most
strongly influenced by broad climatic gradients.

We know from previous research that climate is a significant driver of
productivity across broad spatial scales
\citep{cleveland_relationships_2011}. The majority of studies have
focused on exploring the relationships between productivity and mean
annual temperature (MAT) and mean annual precipitation (MAP), as the
most commonly reported site-level climate variables. These variables
have the advantage that they describe broad trends in temperature and
water availability, and therefore capture a lot of global-scale
variation in climate. There is strong evidence that both MAT and MAP
show significant positive relationships with productivity
\citep{chu_does_2016}. However, as with latitude, the shape of those
relationships is not always clear, and, again, is complicated by the use
of different measures of productivity across studies. There is support
across multiple studies for the hypothesis that various measures of
primary productivity saturate at high levels of MAP, though the
saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
up to 2445mm MAP \citep{schuur_productivity_2003}. Studies of the
influence of MAT on productivity are less conclusive. Luyssaert et al.
\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
while GPP increases linearly with MAT, NPP saturates at around
10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
\citeyearpar{larjavaara_temperature_2012}, find that increases in GPP
saturate at approximately 25\(^\circ\)C MAT, while Schuur
\citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
with temperature. The influence of these climate variables on
productivity is further complicated by the possibility of interactive
effects occurring between them. Taylor et al.
\citeyearpar{taylor_temperature_2017} showed that increased MAP had a
negative effect at low MAT, but a positive effect at high levels of MAT,
and vice versa, suggesting that both temperature and water availability
are important in explaining productivity.

While MAT and MAP are important climate variables, they do not capture
all aspects of climate that may influence productivity, and may be
insufficient to capture the breadth of climatic effects on productivity
\citep{cleveland_relationships_2011}. There is evidence that
productivity also responds to variables such as cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{fyllas_solar_2017}, and potential evapotranspiration
\citep{kerkhoff_plant_2005} in potentially significant ways.
Furthermore, MAT and MAP are very coarse measures of climate, and so
fail to capture much variation in climate on an intra-annual scale,
including the effects of factors such as growing season length, number
of frost-free days, temperature seasonality, and dry season length. Some
studies have suggested that the apparently strong relationship between
MAT and productivity is actually a factor of the correlation between MAT
and growing season length
\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
\citeyearpar{michaletz_convergence_2014} find that, within the growing
season, there is no significant relationship between productivity and
MAT, indicating that the effect of temperature is due to increased
length of growing season, rather than an inherent influence of
temperature on productivity.

In order to approach these broad and complex issues, we simplify the
major gaps in our knowledge to five key questions and corresponding
specific predictions (Table 1). Firstly, we ask how forest autotrophic
carbon fluxes (FACF) vary with latitude. We then test how these fluxes
relate to MAT and MAP, and additionally how they respond to other, less
well studied, climate variables. We finally consider the relationship
between FACF and seasonality, considering the role of seasonality in
explaining variation in carbon fluxes, and the influence of climate on
FACF per month of the growing season.

Resolving these questions requires data which, firstly, control for
variables influencing productivity, including stand age, disturbance
regime, elevation, and methodology; and secondly includes all major
carbon fluxes. Here we use a comprehensive global database of forest
carbon fluxes to explore the above questions for nine carbon fluxes,
allowing for an in-depth exploration of the effect of climate on global
productivity.

@@

\caption{\label{tab:unnamed-chunk-4}Hypothesis table}

@@

Hypotheses \& Specific Predictions & Related references & Overall & GPP & NPP & ANPP & BNPP & \$ANPP\_\{foliage\}\$ & ANPP woody stem & BNPP fine root & R auto root & R auto & Support\\

@@

\multicolumn{1}{l}{\textbf{1. How do FACF vary with latitude?}}\\

@@

\hspace{1em}1.2. Allocation to FACFs varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\

@@

\multicolumn{1}{l}{\textbf{2. How do FACF relate to MAT and MAP?}}\\

@@

\hspace{1em}2.2. FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al. (2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & () & L+ & L+ & yes & Fig. 5\\
\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & I- & L+ & yes & I- & yes & yes & Fig. 4\\

@@

\multicolumn{1}{l}{\textbf{3. How do FACF relate to other annual climate variables?}}\\

@@

\hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & () & yes & \\

@@

\multicolumn{1}{l}{\textbf{4. What is the role of seasonality?}}\\

@@

\hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & () & () & () & () & () & () & () & () & () & \\

@@

\multicolumn{1}{l}{\textbf{5. Per month of the growing seaon, how do FACF vary with climate?}}\\
\hspace{1em}5.1. Increase with temperature (L+) & Michaletz et al. (2014) & (mixed) & () & () & yes & () & yes & () & () & () & () & \\
\hspace{1em}5.2. Increase with PET (L+) &  & (yes) & yes & yes & () & yes & () & yes & yes & () & () & \\
\hspace{1em}5.3. Increase with precipitation (L+) &  & no & () & () & yes & () & yes & () & () & () & () & \\
\hspace{1em}5.4. Increase with solar radiation (L+) &  & (mixed) & yes & yes & () & yes & () & () & yes & () & () & \\

@@

Analyses were conducted on data contained in the open-access ForC
database
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018}.
This database contains records of field-based measurements of forest
carbon stocks and annual fluxes, compiled from original publications and
existing data compilations and databases. Associated data, such as stand
age, measurement methodologies, and disturbance history, are also
included. The database was significantly expanded since the publication
of \citep{anderson-teixeira_forc:_2018} through integration with the
Global Database of Soil Respiration \citep{bond-lamberty_global_2010}.
Additional targeted literature searches were conducted to identify any
further available data on primary productivity, with particular focus on
mature forests in temperate and boreal regions. ForC currently contains
29730 records from 4979 plots, representing 20 distinct ecozones across
all forested biogeographic and climate zones. We used ForC v3.0,
archived on Zenodo with DOI 10.5281/zenodo.3403855.

@@

GPP\strut

@@

NPP\strut

@@

plant tissue; GPP minus ecosystem respiration\strut

@@

ANPP\strut

@@

ANPP\textsubscript{foliage}\strut

@@

ANPP\textsubscript{woody} \textsubscript{stem}\strut

@@

BNPP\textsubscript{root}\strut

@@

BNPP\textsubscript{fine} \textsubscript{root}\strut

@@

R\textsubscript{auto}\strut

@@

R\textsubscript{root}\strut

@@

Measurement of root gas exchange\strut

@@

As discussed above, estimates of NPP, ANPP, and BNPP are generated
through summing measurements of their component parts. Since the
components included in productivity estimates vary between studies,
estimates of productivity were classified within the ForC database
according to their components, and then filtered for analysis. Estimates
of NPP were selected if they included foliage, branch, stem, coarse
root, and fine root. Measures of NPP which included additional
components, including understorey, volatile organic compounds (VOCs),
exudates, estimates of NPP lost to herbivory, and the NPP of
reproductive structures, were excluded. Estimates of ANPP were selected
if they included foliage, stem growth and optionally branch turnover.
Any measures of primary productivity where components were unknown were
excluded from analysis.

@@

\emph{Model specification.} The effects of climate and latitude on
primary productivity were analysed using mixed effects models using the
package `lme4' \citep{bates_fitting_2015} in R v.3.5.1
\citep{r_core_team_r:_2018}. The effect of each extracted climate
variable on each measure of primary productivity was modelled by
specifying the climate variable as a fixed effect. For each climate
variable, three models were specified: a null model; a model with the
climate variable as a linear term; and a model with the climate variable
as a polynomial term. AIC values were calculated for the models and used
to select the best model. If the best model included a polynomial term,
the shape of the polynomial relationship was considered. If the shape of
the relationship \emph{made biological sense}, and was a significant
improvement on the linear relationship (deltaAIC \textgreater{}2), we
accepted the polynomial as the best model. If not, we ran the linear
model as the final model. R\textsuperscript{2} values were calculated
for the best model. All R\textsuperscript{2} values presented here are
marginal R\textsuperscript{2} values, and refer to the proportion of
variation explained by only the fixed effects, unless otherwise
specified. In addition, slope coefficients were calculated for the
linear models.

@@

R\textsuperscript{2} values were generally highest in the major fluxes,
and decreased in subsidiary fluxes (Supporting Information S2). Of the
major fluxes, R\textsubscript{auto} and GPP were the most strongly
explained by latitude and climate, with climate explaining at most 71\%
of variation in GPP, and 65\% in R\textsubscript{auto}. The proportion
of variation explained by climate and latitude decreased in NPP and
ANPP, with climate explaining at most 51\% of variation in NPP and 44\%
in ANPP. Of the major fluxes, BNPP\textsubscript{root} was the least
well explained by climate and latitude, with climate explaining at most
36\% of variation.

With the exception of ANPP\textsubscript{foliage}, the proportion of
variation explained by climate and latitude in subsidiary fluxes was
much lower. Climate explained at most 24\% of variation in
ANPP\textsubscript{woody} \textsubscript{stem}, 19\% in
BNPP\textsubscript{fine} \textsubscript{root}, and 27\% in
R\textsubscript{root}. In contrast, climate strongly explained variation
in ANPP\textsubscript{foliage}, with mean annual temperature explaining
58\% of variation. This pattern was also seen in the
R\textsuperscript{2} values for multivariate models.

\emph{How does productivity relate to MAT and MAP?} We focus first on
considering the relationship between productivity and MAT and MAP. MAT
and MAP are the most commonly reported site-level climate variables, and
much previous research into the effect of climate on forest productivity
has focused on these as key climate variables. MAT was a significant
(p\textless{}0.05) and strong predictor of productivity for all carbon
fluxes tested, with all fluxes showing a linear increase with
temperature (fig. 5). We found no support for a saturation point of
productivity with temperature.

MAP was found to be a significant (p\textless{}0.05) but poor predictor
of productivity, explaining, with the exception of
R\textsubscript{auto}, at most 37\% of variation in carbon flux. For the
majority of fluxes productivity was best predicted by a polynomial
model. Productivity increased with precipitation, up until a saturation
point at between 3000 and 4000mm annual precipitation, above which
productivity started to decrease (fig. 5). The notable exception to this
was GPP: the model indicated that GPP continued to increase with
precipitation up to measures of at least 5000mm annually
(p\textless{}0.0001, R\textsuperscript{2} = 0.33. Data above this point
was not available, but the model trend indicated that the saturation
point for this model would be around 5000mm MAP.

There was a significant interactive effect between MAT and MAP for GPP,
BNPP\textsubscript{root}, BNPP\textsubscript{fine} \textsubscript{root},
ANPP, ANPP\textsubscript{woody} \textsubscript{stem}, and
R\textsubscript{root} (fig.4). There was a significant additive effect
for R\textsubscript{auto}. NPP and ANPP\textsubscript{foliage} showed no
significant interactive or additive effect: including MAP as a second
variable did not improve on the model including only MAT.

@@

We found a significant relationship between productivity and potential
evapotranspiration for all fluxes. ANPP\textsubscript{foliage},
BNPP\textsubscript{fine} \textsubscript{root} and R\textsubscript{root}
increased linearly with PET; however, all other fluxes showed a
polynomial relationship with PET (fig. 5). We found strong evidence for
a saturation point or peak with PET: productivity tended to increase at
values below 1000mm, before saturating between 1200 and 1700mm. There
was evidence that productivity begins to decrease at values above 1800mm
PET.

Vapour pressure deficit was a significant predictor of productivity for
all fluxes. BNPP\textsubscript{fine} \textsubscript{root} showed a
linear relationship with vapour pressure deficit (R\textsuperscript{2} =
0.07, p\textless{}0.05), but all other fluxes showed a polynomial
relationship (fig. 5). Productivity initially increased with vapour
pressure deficit, before saturating at around 0.8 kPa. At values above
0.8 kPa, productivity began to decrease.

>>>>

{[}\textbf{very rough start:}{]} \ldots{}.. Here, we drew upon \#
records from the Global Forest Carbon Database (ForC), representing all
major forest types and the nine most significant forest autotrophic
carbon flux (FACF) variables, to comprehensively explore how C cycling
in mature, undisturbed forests varies with latitude and climate on a
global scale. We show that, across all FACF variables analyzed, C
cycling decreases linearly with latitude --a finding that confirms
multiple previous studies but contradicts the idea that productivity of
temperate forests rivals that of tropical forests. The FACF variables
increase in proportion to one another, with no differences in allocation
detectable at this global scale, but differed in that latitude explained
a lower proportion of variation among subsidiary fluxes (in particular,
woody aboveground productivity and belowground net primary productivity,
\(BNPP\)). Climate explained a significant proportion (\#-\#\%) of
variation in all C fluxes analyzed (less for subsidiary fluxes), with
temperature variables in general and mean annual temperature (\(MAT\))
in particular being the best predictors of FACF on this global scale.
While other climate variables (\emph{e.g.}, displayed significant
correlation with FACF, none of them improved significantly over the
explanatory power \(MAT\). The effects of temperature were modified by
moisture availability, with reduced FACF under hot and dry conditions
and sometimes under very high precipitaiton (especially for \(BNPP\)).
FACF declined with temperature seasonality, but growing season length
doesn't improve upon MAT as a predictor. Within the growing season, the
influence of climate on C cycling is smaller but still significant for a
number of carbon fluxes. These findings clarify the big picture of how
FACF varies with latitude and climate on a global scale. As we enter a
period of accelerating climatic change, understanding of the fundamental
climatic controls on FACF sets a foundation for understanding patterns
of change.

@@

\textbf{Forests play a critical role in regulating atmospheric CO\_2 and
climate} (\textbf{Bonan}). The total amount of CO\_2 cycling through
Earth's forests each year (\emph{i.e.}, total gross primary production,
\(GPP\)) is more than five times anthropogenic fossil fuel emissions,
and net sequestration (\textasciitilde{}2.4 GtCyr\textsuperscript{-1})
offsetting roughly 30\% of these emissions \citep{pan_large_2011}. As
atmospheric carbon dioxide levels continue to rise, with consequences
for global climate, there is increasing recognition that proper
protection and management of forest resources will have an important
role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and
land reports**). Carbon fixation, allocation, and metabolism by trees
set the basis for energy and material flows in forest ecosystems and
define their interactions with Earth's changing climate.

\textbf{Forests' autotrophic carbon flux (FACF, bbl:define) influences
all organic matter stocks in ecosystems and is linked to cycling of
energy, water, and nutrients} (\textbf{REFS; \emph{Piao et al. 2010.
Forest annual carbon cost: a global‐scale analysis of autotrophic
respiration. Ecology 91(3), 652-661Ƴ} }). Forest productivity sets the
energy ultimately available to heterotrophic organisms (including
microbes), in turn influencing their abundance (\textbf{REFS}) and
possibly diversity
(\href{https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257}{Waide
et al. 2009}\textbf{REFS})--but probably not individual metabolic rate
(\textbf{Anderson \& Jetz 2005, but check more recent literature}).
However, our ability to draw general macroscopic conclusions regarding
global variation in multiple productivity variables with respect to
climate has been limited in that these analyses often mix forests that
vary in stand age, disturbance history, and/or management status; do not
always sufficiently parse related variables (\emph{e.g.}, combining net
primary productivity records with and without belowground components);
and typically consider only one or a few variables at a time.

\textbf{The productivity of forests varies with latitude}, showing a
general trend of decreasing productivity with latitude
\citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
productivity is lowest in the boreal regions, and increases into the
temperate regions
\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
However, evidence is inconclusive on whether productivity continues to
increase into the tropics, or whether it plateaus in temperate regions.
Evidence for this is further complicated by the fact that different
studies use different measures of productivity to explore these
relationships. For example, modelling of global terrestrial ecosystem
gross primary productivity (GPP) through upscaling and calibration of
eddy flux measurements indicates that GPP peaks in tropical forests
\citep{beer_terrestrial_2010, jung_global_2011};\href{https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729}{Bagdley
et al. 2019}). This is corroborated by analysis of site-level GPP
measurements, which appear to reach their highest levels in tropical
forests \citep{luyssaert_co_2007}. In contrast, there is evidence that
the highest values of net primary productivity (NPP) may be found in
temperate forests \citep{luyssaert_co_2007, huston_global_2009},
although other studies find NPP is highest in the tropics, showing a
decrease with latitude \citep{simova_enigma_2017}. Other studies have
chosen to focus exclusively on above-ground net primary productivity
(ANPP), finding evidence of a weak negative relationship between ANPP
and latitude \citep{huston_global_2009, gillman_latitude_2015}.

\textbf{Climate is a significant driver of productivity across broad
spatial scales} \citep{cleveland_relationships_2011}. The majority of
studies have focused on exploring the relationships between productivity
and mean annual temperature (MAT) and mean annual precipitation (MAP),
as the most commonly reported site-level climate variables. These
variables have the advantage that they describe broad trends in
temperature and water availability, and therefore capture a lot of
global-scale variation in climate. There is strong evidence that both
MAT and MAP show significant positive relationships with productivity
\citep{chu_does_2016}. However, as with latitude, the shape of those
relationships is not always clear, and, again, is complicated by the use
of different measures of productivity across studies. Various measures
of primary productivity saturate at high levels of MAP, though the
saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
up to 2445mm MAP \citep{schuur_productivity_2003}. Studies of the
influence of MAT on productivity are less conclusive. Luyssaert et al.
\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
while GPP increases linearly with MAT, NPP saturates at around
10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
\citeyearpar{larjavaara_temperature_2012}, find that increases in GPP
saturate at approximately 25\(^\circ\)C MAT, while Schuur
\citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
with temperature. bbl:maybe remove following sentences The influence of
these climate variables on productivity is further complicated by the
possibility of interactive effects occurring between them. Taylor et al.
\citeyearpar{taylor_temperature_2017} showed that increased MAP had a
negative effect at low MAT, but a positive effect at high levels of MAT,
and vice versa, suggesting that both temperature and water availability
are important in explaining productivity.

\textbf{Primary productivity can be influenced by many other factors as
well}, which often act across a range of scales, and may show
interactive effects with each other
\citep{cleveland_relationships_2011}. On a local scale, stand age
\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity (Liang
et al. 2019. Positive biodiversity-productivity relationship predominant
in global forests. Science 354, Issue 6309, aaf8957), management
\citep{simova_enigma_2017}; nutrient availability
\citep{aragao_above-_2009}; and altitude
\citep{girardin_net_2010, malhi_variation_2017} all impact forest
productivity. On a global scale, we expect that productivity is most
strongly influenced by broad climatic gradients. There is evidence that
productivity also responds to variables such as cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{fyllas_solar_2017}, and potential evapotranspiration
\citep{kerkhoff_plant_2005} in potentially significant ways.
Furthermore, MAT and MAP are very coarse measures of climate, and so
fail to capture much variation in climate on an intra-annual scale,
including the effects of factors such as growing season length, number
of frost-free days, temperature seasonality, and dry season length. Some
studies have suggested that the apparently strong relationship between
MAT and productivity is actually a factor of the correlation between MAT
and growing season length
\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
\citeyearpar{michaletz_convergence_2014} find that, within the growing
season, there is no significant relationship between productivity and
MAT, indicating that the effect of temperature is due to increased
length of growing season, rather than an inherent influence of
temperature on productivity.

\textbf{The recent development of a global forest carbon database
synthesizing multiple variables and including records of stand history
(ForC;
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
opens up the possibility for a standardized analysis of global scale
variation} in multiple components of forest productivity and the
principle climatic drivers of these patterns. In order to approach these
broad and complex issues, we simplify the major gaps in our knowledge to
five broad hypotheses and corresponding specific predictions (Table 1).
First, we ask how forest autotrophic carbon fluxes (FACF) vary with
latitude. We then test how these fluxes relate to MAT and MAP, and
additionally how they respond to other, less well studied, climate
variables. We finally consider the relationship between FACF and
seasonality, considering the role of seasonality in explaining variation
in carbon fluxes, and the influence of climate on FACF per month of the
growing season. We use a comprehensive global database of forest carbon
fluxes to explore the above questions for nine carbon fluxes, allowing
for an in-depth exploration of the effect of climate on global
productivity.

@@

\caption{\label{tab:unnamed-chunk-4}**Summary of hypotheses, corresponding specific predictions, and results.** Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD). 'n.s.' indicates that the relationship is not signficant.}

@@

Hypotheses \& Specific Predictions & Related references & Overall & GPP & NPP & ANPP & BNPP & ANPP foliage & ANPP woody stem & BNPP fine root & R auto root & R auto & Support\\

@@

\multicolumn{1}{l}{\textbf{H1. FACFs decrease linearly with latitude, and in constant proportion.}}\\

@@

\hspace{1em}1.2. Allocation of GPP to subsidiary fluxes varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\

@@

\multicolumn{1}{l}{\textbf{H2. FACFs increase with MAT, with an interactive effect of MAP.}}\\

@@

\hspace{1em}2.2. FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al. (2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & n.s. & L+ & L+ & yes & Fig. 5\\
\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 4\\

@@

\multicolumn{1}{l}{\textbf{H3. FACFs are strongly correlated with other annual climate variables.}}\\

@@

\hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & n.s. & yes & \\

@@

\multicolumn{1}{l}{\textbf{H4. FACFs are reduced under seasonal climates.}}\\

@@

\hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & \\

@@

\multicolumn{1}{l}{\textbf{H5. Considering only growing season months, FACF correlates with climatic drivers.}}\\
\hspace{1em}5.1. Increase with temperature (L+) & Michaletz et al. (2014) & (mixed) & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
\hspace{1em}5.2. Increase with PET (L+) &  & (yes) & yes & yes & n.s. & yes & n.s. & yes & yes & n.s. & n.s. & \\
\hspace{1em}5.3. Increase with precipitation (L+) &  & no & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
\hspace{1em}5.4. Increase with solar radiation (L+) &  & (mixed) & yes & yes & n.s. & yes & n.s. & n.s. & yes & n.s. & n.s. & \\

@@

Analyses were conducted on data contained in the open-access ForC
database
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018}.
This database contains records of field-based measurements of forest
carbon stocks and annual fluxes, compiled from original publications and
existing data compilations and databases. Associated data, such as stand
age, measurement methodologies, and disturbance history, are also
included. The database was significantly expanded since the publication
of \citep{anderson-teixeira_forc:_2018} through integration with the
Global Database of Soil Respiration Database
\citep{bond-lamberty_global_2010}. Additional targeted literature
searches were conducted to identify any further available data on
primary productivity, with particular focus on mature forests in
temperate and boreal regions. ForC currently contains 29730 records from
4979 plots, representing 20 distinct ecozones across all forested
biogeographic and climate zones. We used ForC v3.0, archived on Zenodo
with DOI 10.5281/zenodo.3403855.

@@

\(GPP\)\strut

@@

\(NPP\)\strut

@@

plant tissue; \(GPP - R_{auto}\)\strut

@@

\(ANPP\)\strut

@@

\(ANPP_{foliage}\)\strut

@@

\(ANPP_{woody-stem}\)\strut

@@

\(BNPP_{root}\)\strut

@@

\(BNPP_{fine.root}\)\strut

@@

\(R_{auto}\)\strut

@@

\(R_{root}\)\strut

@@

\emph{Measurement of root gas exchange, root exclusion from soil
respiration chambers}\strut

@@

As discussed above, estimates of \(NPP\), \(ANPP\), and \(BNPP\) are
generated through summing measurements of their component parts. Since
the components included in productivity estimates vary between studies,
estimates of productivity were classified within the ForC database
according to their components, and then filtered for analysis. Estimates
of NPP were selected if they included foliage, branch, stem, coarse
root, and fine root. Measures of NPP which included additional
components, including understorey, volatile organic compounds (VOCs),
exudates, estimates of NPP lost to herbivory, and the NPP of
reproductive structures, were excluded. Estimates of ANPP were selected
if they included foliage, stem growth and optionally branch turnover.
Any measures of primary productivity where components were unknown were
excluded from analysis.

@@

\emph{Model specification.} The effects of climate and latitude on
primary productivity were analysed using mixed effects models using the
package `lme4' \citep{bates_fitting_2015} in R v.3.5.1
\citep{r_core_team_r:_2018}. The effect of each extracted climate
variable on each measure of primary productivity was modelled by
specifying the climate variable as a fixed effect. For each climate
variable, three models were specified: a null model; a model with the
climate variable as a linear term; and a model with the climate variable
as a polynomial term. AIC values were calculated for the models and used
to select the best model. If the best model included a polynomial term,
the shape of the polynomial relationship was considered. If the shape of
the relationship \emph{made biological sense}, and was a significant
improvement on the linear relationship (deltaAIC \textgreater{}2), we
accepted the polynomial as the best model. If not, we ran the linear
model as the final model. \(R^2\) values were calculated for the best
model. All \(R^2\) values presented here are marginal \(R^2\) values,
and refer to the proportion of variation explained by only the fixed
effects, unless otherwise specified. In addition, slope coefficients
were calculated for the linear models.

@@

\(R^2\) values were generally highest in the major fluxes, and decreased
in subsidiary fluxes (Supporting Information S2). Of the major fluxes,
R\textsubscript{auto} and GPP were the most strongly explained by
latitude and climate, with climate explaining at most 71\% of variation
in GPP, and 65\% in R\textsubscript{auto}. The proportion of variation
explained by climate and latitude decreased in NPP and ANPP, with
climate explaining at most 51\% of variation in NPP and 44\% in ANPP. Of
the major fluxes, BNPP\textsubscript{root} was the least well explained
by climate and latitude, with climate explaining at most 36\% of
variation.

With the exception of ANPP\textsubscript{foliage}, the proportion of
variation explained by climate and latitude in subsidiary fluxes was
much lower. Climate explained at most 24\% of the variation in
ANPP\textsubscript{woody} \textsubscript{stem}, 19\% in
BNPP\textsubscript{fine} \textsubscript{root}, and 27\% in
R\textsubscript{root}. In contrast, climate strongly explained variation
in ANPP\textsubscript{foliage}, with mean annual temperature explaining
58\% of variation. This pattern was also seen in the \(R^2\) values for
multivariate models.

\emph{How does productivity relate to MAT and MAP?} MAT and MAP are the
most commonly reported site-level climate variables, and much previous
research into the effect of climate on forest productivity has focused
on these as key climate variables. MAT was a significant
(p\textless{}0.05) and strong predictor of productivity for all carbon
fluxes tested, with all fluxes showing a linear increase with
temperature (fig. 5). We found no support for a saturation point of
productivity with temperature.

MAP was found to be a significant (p\textless{}0.05) but poor predictor
of productivity, explaining, with the exception of
R\textsubscript{auto}, at most 37\% of variation in carbon flux. For the
majority of fluxes productivity was best predicted by a polynomial
model. Productivity increased with precipitation, up until a saturation
point at between 3000 and 4000mm annual precipitation, above which
productivity started to decrease (fig. 5). The notable exception to this
was GPP: the model indicated that GPP continued to increase with
precipitation up to measures of at least 5000mm annually
(p\textless{}0.0001, \(R^2\) = 0.33. Data above this point was not
available, but the model trend indicated that the saturation point for
this model would be around 5000mm MAP.

There was a significant interactive effect between MAT and MAP for GPP,
BNPP\textsubscript{root}, BNPP\textsubscript{fine} \textsubscript{root},
ANPP, ANPP\textsubscript{woody} \textsubscript{stem}, and
R\textsubscript{root} (fig.4). There was a significant additive effect
for R\textsubscript{auto}. NPP and ANPP\textsubscript{foliage} showed no
significant interactive or additive effect.

@@

We found a significant relationship between productivity and potential
evapotranspiration for all fluxes. ANPP\textsubscript{foliage},
BNPP\textsubscript{fine} \textsubscript{root} and R\textsubscript{root}
increased linearly with PET; however, all other fluxes showed a
polynomial relationship with PET (fig. 5). We found strong evidence for
a saturation point or peak with PET: productivity tended to increase at
values below 1000mm, before saturating between 1200 and 1700mm. There
was also evidence that productivity begins to decrease at values above
1800mm PET.

Vapour pressure deficit was a significant predictor of productivity for
all fluxes. BNPP\textsubscript{fine} \textsubscript{root} showed a
linear relationship with vapour pressure deficit (\(R^2\) = 0.07,
p\textless{}0.05), but all other fluxes showed a polynomial relationship
(fig. 5). Productivity initially increased with vapour pressure deficit,
before saturating at around 0.8 kPa. At values above 0.8 kPa,
productivity began to decrease.

All fluxes, with the exception of R\textsubscript{root}, showed a
positive linear relationship with solar radiation. Solar radiation
explained a low proportion of variability in productivity for all
fluxes, explaining less than 20\% of the variation in each flux, with
the exception of R\textsubscript{auto} (\(R^2\) = 0.26,
p\textless{}0.05).
